%-------------------------
% Resume in Latex (adapted from template.tex)
%------------------------

\documentclass[letterpaper,10pt]{extarticle}

\usepackage{latexsym}
\usepackage[margin=0.62in]{geometry}
\usepackage{titlesec}
\usepackage[usenames,dvipsnames]{color}
\usepackage{enumitem}
\usepackage[pdftex]{hyperref}
\usepackage{fancyhdr}

\pagestyle{fancy}
\fancyhf{}
\fancyfoot{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

% Adjust margins
\urlstyle{same}
\raggedbottom
\raggedright
\setlength{\tabcolsep}{0in}

% Sections formatting
\titleformat{\section}{
  \vspace{-4pt}\scshape\raggedright\normalsize
}{}{0em}{}[\color{black}\titlerule \vspace{-5pt}]

% Custom commands
\newcommand{\resumeItem}[2]{
  \item\small{\textbf{#1}: #2 \vspace{-2pt}}
}

\newcommand{\resumeSubheading}[4]{
  \vspace{-1pt}\item
    \begin{tabular*}{0.97\textwidth}{l@{\extracolsep{\fill}}r}
      \textbf{#1} \textit{\small / #3 / #4} & \textit{\small #2} \\
    \end{tabular*}\vspace{-5pt}
}

\newcommand{\resumeSubheadingCompact}[4]{
  \vspace{-1pt}\item
    \begin{tabular*}{0.97\textwidth}{l@{\extracolsep{\fill}}r}
      \textbf{#1} \textit{\small / #3 / #4} & \textit{\small #2} \\
    \end{tabular*}\vspace{-9pt}
}

\newcommand{\resumeSubItem}[2]{\resumeItem{#1}{#2}\vspace{-4pt}}

\renewcommand{\labelitemii}{$\circ$}

\newcommand{\resumeSubHeadingListStart}{\begin{itemize}[leftmargin=*]}
\newcommand{\resumeCompactSectionListStart}{\begin{itemize}[leftmargin=*, itemsep=0pt, topsep=2pt, parsep=0pt, partopsep=0pt]}
\newcommand{\resumeEducationListStart}{\begin{itemize}[leftmargin=*, itemsep=-1pt, topsep=1pt, parsep=0pt, partopsep=0pt]}
\newcommand{\resumeSubHeadingListEnd}{\end{itemize}}
\newcommand{\resumeItemListStart}{\begin{itemize}[leftmargin=*, itemsep=0pt, topsep=2pt, parsep=0pt, partopsep=0pt]}
\newcommand{\resumeItemListEnd}{\end{itemize}\vspace{-4pt}}

\begin{document}

%----------HEADING-----------------
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
  \textbf{\Large Pranshu Gupta} & Email: \href{mailto:pranshug258@outlook.com}{pranshug258@outlook.com} \\
  \href{https://pranshug.com}{https://pranshug.com} $\vert$ \href{https://linkedin.com/in/pranshug258/}{linkedin.com/in/pranshug258} & Mobile: +1 (425) 679-2402 \\
\end{tabular*}

%-----------EXPERIENCE-----------------
\section{Professional Experience}
\resumeSubHeadingListStart

  \resumeSubheading
    {Microsoft AI}{Aug 2025 -- Present}
    {Member of Technical Staff}{Redmond, WA}
    \resumeItemListStart
      \item\small{Serve as \textbf{technical area owner} for LLM inference on Copilot, shaping architectural decisions across model integration, safety infrastructure, and inference reliability.}
      \item\small{Engineered the inference pipeline for \textbf{GPT-5.1 instant and reasoning models} on Copilot at launch, with meta-prompting for personality alignment, delivering a \textbf{30\% improvement in personality evals} for the reasoning path.}
      \item\small{Implemented a new \textbf{constrained sampling approach} for tool calls in GPT-5.1 reasoning models, verified through token-level analysis, preventing a production regression in tool call reliability at Copilot launch.}
      \item\small{Scoped, designed, and built \textbf{GPU-based safety classifier} support in Chat Completions API: non-streaming endpoints returning classification logits for client-side threshold scoring across input and output streams.}
      \item\small{Designed and shipped an \textbf{incremental CoT summarization} system: buffers streaming reasoning chunks and progressively surfaces summaries in the UI, keeping it responsive while protecting raw model reasoning as IP.}
      \item\small{Reduced inference costs by eliminating non-essential classifiers (\textbf{reducing GPU footprint by 20\%}) and resolved a token overhead inefficiency in Chat Completions API (\textbf{reducing input tokens by 25\%} and \textbf{TTFT by 30ms}).}
      \item\small{Defined and instrumented LLM API \textbf{traffic classification} dimensions (product, client app, inference intent, license tier) across MAI and M365 Copilot; built Grafana dashboards informing capacity planning and cost attribution across both orgs.}
    \resumeItemListEnd

  \resumeSubheading
    {Microsoft Azure AI}{Jan 2025 -- Aug 2025}
    {Senior Software Engineer}{Redmond, WA}
    \resumeItemListStart
      \item\small{Designed and built a distributed \textbf{token-based rate limiter} using Tiktoken for precise token counting, replacing a length-based estimator with 219\% average error; implemented pub-sub quota state synchronization across instances.}
      \item\small{Designed and built the inference pipeline for an intelligent \textbf{model router}: injecting per-request model selection based on cost/complexity signals before each LLM call, reducing per-inference cost by \textbf{55\%}.}
    \resumeItemListEnd

  \resumeSubheading
    {Microsoft Azure Core}{Mar 2024 -- Jan 2025}
    {Senior Software Engineer}{Redmond, WA}
    \resumeItemListStart
      \item\small{Served as \textbf{technical area owner} for resource lifecycle management in Azure Resource Manager.}
      \item\small{Mentored a new-grad engineer through weekly 1:1s; they now independently own a critical component in control plane.}
      \item\small{Designed and shipped a multi-bucket \textbf{token-based background job throttling} system with retry-storm protection, preventing noisy-neighbor throttling in Azure Control Plane background-job infrastructure.}
      \item\small{Single-handedly designed and shipped \textbf{async operation URI signing}; ran office hours with resource provider teams to guide API surface changes, adopted by \textbf{242 service teams} with \textbf{220+ enforcing}.}
    \resumeItemListEnd

  \resumeSubheading
    {Microsoft Azure Core}{Jun 2020 -- Mar 2024}
    {Software Engineer II}{Redmond, WA}
    \resumeItemListStart
      \item\small{Designed and shipped \textbf{async operation callbacks} replacing polling, cutting control plane polling traffic by \textbf{94\%}.}
      \item\small{Led \textbf{Azure region teardown} design and implementation end-to-end: halted new resource creation, resolved subscription/resource group hierarchies, and migrated live resources to healthy regions with zero data loss.}
      \item\small{Designed and built a \textbf{health-driven deployment pipeline} for ARM control-plane manifests: staged multi-region rollouts with post-step RP health checks that halt or rollback on degradation.}
    \resumeItemListEnd

  \resumeSubheading
    {Microsoft Core Services}{Jun 2017 -- Aug 2019}
    {Software Engineer}{Hyderabad, India}
    \resumeItemListStart
      \item\small{Modernized a data enrichment platform to Azure Function Apps (\textbf{\$4K/month} savings); integrated Cognitive Services for data standardization, and built a bulk-processing pipeline with \textbf{40--50x throughput} improvement.}
    \resumeItemListEnd

\resumeSubHeadingListEnd

%-----------SKILLS-----------------
\section{Technical Skills}
\begin{itemize}[leftmargin=*, itemsep=0pt, topsep=2pt, parsep=0pt, partopsep=0pt]
  \item\small{\textbf{AI/ML \& Inference}: LLM inference/serving, prompt engineering, model evaluations, safety orchestration, RAG}
  \item\small{\textbf{Languages \& Systems}: Python, .NET, C++, KQL, JavaScript, HTML, CSS, Kubernetes, Redis, Grafana}
  \item\small{\textbf{Cloud \& Infrastructure}: Azure (Cosmos DB, ARM Templates, Service Bus, DevOps), AWS (Bedrock, S3, ECR)}
\end{itemize}

%-----------PUBLICATIONS-----------------
\section{Publications \& Writing}
\resumeCompactSectionListStart
  \item\small{\href{https://pranshug.com/blog/runtime-optimizations-for-llms}{\textit{Runtime Optimizations for LLMs}} (2026). Production notes on inference optimizations for latency/cost-efficient LLM serving.}
  \item\small{\href{https://dl.acm.org/doi/10.1145/3613904.3642078}{\textit{The Observer Effect on Social Media}} — ACM CHI (2024). Causal inference and time series study on longitudinal Facebook data from 300+ participants to quantify how monitoring awareness alters online behavior.}
  \item\small{\href{https://www.nature.com/articles/s41599-021-00993-6}{\textit{Mental Health Impact of Active-Shooter Drills}} — Nature (2021). Applied ML classifiers and interrupted time series analysis to 54M social media posts across 114 schools to measure psychological and civic impact.}
\resumeSubHeadingListEnd

%-----------VOLUNTEER-----------------
\section{Volunteer Work}
\resumeCompactSectionListStart
  \item\small{\textbf{AI Lab Assistant} (TurnTheBus nonprofit, 2025--Present): Architected and built end-to-end on \textbf{AWS} using \textbf{RAG} over physics lab manuals to guide high school students through virtual physics experiments.}
\resumeSubHeadingListEnd

%-----------EDUCATION-----------------
\section{Education}
\resumeEducationListStart
  \resumeSubheadingCompact
    {Georgia Institute of Technology}{May 2020}
    {M.S. Computer Science (Machine Learning)}{Atlanta, GA}
  \resumeSubheadingCompact
    {Indian Institute of Technology, Kanpur}{May 2017}
    {B.Tech, C.S. and Eng.}{Kanpur, India}
\resumeSubHeadingListEnd

\end{document}
